---
title: "timecoursedata: Time course gene expression datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
    %\VignetteIndexEntry{Installation}
      %\usepackage[UTF-8]{inputenc}
---

## Overview

This package provides time course gene expression datasets. The data included
in the package and its processing is described below. 

Overviw of the datasets included in timecoursedata:

object | description
------- | ------------
`shoemaker2015` | microarray data from influenza-exposed mice
`varoquaux2019leaf` | RNA-seq from leaf samples of *S. bicolor* exposed to drought
`varoquaux2019root` | RNA-seq from root samples of *S. bicolor* exposed to drought



## Data structure

Each datasets in this package are represented by:

- a matrix containing the gene expression data
- a data.frame containing the metadata associated with it

The metadata contains at least the following information:

column | description
----- | ----------
`row.names` | name of the sample
`Timepoint` | which timepoints this sample belongs to
`Group` | which group this sample belongs to
`Replicate` (optional) | which replicate this sample belongs to 


## An Ultrasensitive Mechanism Regulates Influenza Virus-Induced Inflammation

This dataset is a microarray time-course experiment, exposing mice to three
different strains of influenza with varying doses. The authors of the
experiment collected three replicates of lung tissue samples during 14
unevenly-spaced time-points after infection, resulting in a dataset of 209
samples.

```{r}
library(timecoursedata)
data(shoemaker2015)
head(shoemaker2015$meta)
head(shoemaker2015$data)
```

## Transcriptomic analysis of field-droughted sorghum from seedling to maturity reveals biotic and metabolic responses

This pairs of datasets (`varoquaux2019leaf` and `varoquaux2019root`)
corresponds to transcriptomic data from respectively leaf and root samples of
*S. bicolor* exposed to droughts (pre-flowering and post-flowering drougt)
from seedling to maturity. Samples were collected weekly during 15 weeks. 
The data provided here is not normalized.

```{r}
data(varoquaux2019leaf)
nrow(varoquaux2019leaf$data)
ncol(varoquaux2019leaf$data)

```
